---
title: "Data Manupulation"
output: github_document 
---

```{r set up, include = FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(tidyverse)
library(readxl)

``` 

## import datasets
```{r}
litters_data =  read_csv("./data_import_examples/FAS_litters.csv")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data_import_examples/FAS_pups.csv")
pups_data = janitor::clean_names(pups_data)

```

## selecting !!! 

```{r}
select(litters_data,group,litter_number)
select(litters_data,group,litter_number,gd0_weight)
select(litters_data,group,gd0_weight, starts_with("pups"))
select(litters_data, litter_number,group,gd0_weight)
select(litters_data,gd0_weight,everything())
select(litters_data,-group)
select(litters_data,group,litter_number,gd0_weight:gd_of_birth)
select(litters_data,GROUP = group, litter_number)

rename(litters_data,GROUP = group)
```

***sequencing; renaming the variable; detemrine the first and last variable
The difference between 35 and 33: 35 is only changing the variable name and show all of variables. 35 not only changed the variable name, but also choose the variables to be shown. ***

## filtering !!!
```{r}
filter(litters_data, group == "Mod8")
filter(litters_data, gd_of_birth == 20)
filter(litters_data, gd_of_birth < 20)
filter(litters_data, gd_of_birth >= 20)
filter(litters_data,gd_of_birth < 6, group == "Con7")
# following two lines produce same results##
filter(litters_data,group == "Con7" | group == "Mod8")
filter(litters_data,group %in% c("Con7","Mod8"))
filter(litters_data,gd0_weight + gd18_weight < 70)
#getting the missing value, but don't do this please when we return the dataset, they only have missing values.#
filter(litters_data, is.na(gd0_weight))
# what we did instead: dropping the missing values under gd0_weight"# 
drop_na(litters_data,gd0_weight)

```
*** "==" is to asking question, whether the value is the specific thing in the dataset.
"=" is assigning a value to variable.
we can put multiple conditions together. 
! --> not, can be put outside of logical condition. 
%in% --> check more than one groups

## Mutate !!!
```{r}
mutate(
  litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```
???why we need to make group lower case? 

## Arrange 
```{r}
arrange(litters_data, pups_born_alive)
arrange(litters_data, desc(pups_born_alive))
arrange(litters_data, pups_born_alive,gd0_weight)
```








